---
title: "Covid-19 Death Investigation"
author: "wliu16"
date: "2022-12-14"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(caret)
library(ROCR)
library(ggplot2)
library(pROC)
library(rpart)
library(rpart.plot)
library(cutpointr)
library(ROSE)
```

```{r}
setwd("~/Desktop/R/64060-003/Final_Exam")  #set working directory
Covid <- read.csv("Covid Data.csv")  #load the data
summary(Covid) 
```

**Data Preparation**<br>  

```{r}
#Since all the boolean in the dataset uses 1("Yes") and 2("No"), convert rest of the columns

#Convert non-Death cases to NA
Covid$DATE_DIED <- as.Date(Covid$DATE_DIED)
#Convert values of 97, 98 and 99 to NAs in all binary columns 
Covid <- Covid %>% na_if(97) %>% na_if(98) %>% na_if(99)

#Convert classification from 1-7 to binary 1("Yes") and 2("No")
Covid <- Covid%>%
  mutate(CLASSIFICATION = ifelse(CLASIFFICATION_FINAL<=3, 1 , 2))%>%
  filter(CLASSIFICATION == 1) #keep only covid-positive cases
  
#Convert death from date to binary 1("Yes") and 2("No")
Covid <- Covid%>%
  mutate(DEATH = ifelse(is.na(DATE_DIED), 2, 1))

#391979 obs of 23 variables
```

**Data Exploration**
```{r}
xtabs(~DEATH, data = Covid)
xtabs(~DEATH+SEX, data = Covid) #death and covid distribution by sex
xtabs(~DEATH+PREGNANT, data = Covid) #death and covid distribution by PREGNANCY
```

**Comment**<br>  
Early observations for death and covid cases: <br>  
1) Male (sex =2) death rate is higher than female whether is covid positive or not<br>  
2) Since prenant women is only a small portion of female (1.55%) and our study
is not focused on pregnant women, let's delete pregnant variable for the study
3) We can't impute the 345 NA records for age. Let's remove it since it is a 
small portion. <br>  
4) Let's also remove USMER, MEDICAL_UNIT. Not important where patients receive care<br>  
5) INTUBED and ICU still have high ratio of NAs. Assumne only severe cases need
those procedures, it's ok to replace those NA with median which is 2("no") <br>  
6) The rest of NAs usually count than less than 1% of column data so let's impute 
the NAs by median number<br>  

<br>  

**Data Preparation**<br>  

```{r}
Covid <- Covid[-c(3, 6, 17, 18, 21, 22)] #delete pregnant, USMER, MEDICAL_UNIT etc
#391979 obs. of  17 variables
Covid <- Covid%>%
  filter(!is.na(AGE))  #391853 obs. of  17 variables

#impute missing values with mean
Covid[, c(4:16)] <- Covid[, c(4:16)]%>%
    mutate_if(is.numeric, function(x) ifelse(is.na(x), median(x, na.rm = T), x))

#Change data attribute from character to factor, the data is coded as 1 as no and 2 as yes
Covid$DEATH <- as.factor(Covid$DEATH) 
```

**Data partition**
```{r}
#Partition the given training data into 70% training data and 30% testing data
set.seed(100)
index_train <- createDataPartition(Covid$DEATH, p=0.7, list= F)
Covid_train <- Covid[index_train, ]
Covid_test <- Covid[-index_train, ]
```
<br>  

**Run logistic regression model**
```{r}
set.seed(1)
log_model <- glm(DEATH~., data = Covid_train, family = 'binomial')
log_model
```
**Comment**<br>  
Factors will increase the chance of death: <br>  
SEX(male), AGE (high), PATIENT TYPE (hospitalized), PNEUMONIA (positive), 
DIABETES (positive), COPD (negative), ASTHMA (negative), INMSUPR (positive),
HIPERTENSION (positive), CARDIOVASCULAR (negative), RENAL_CHRONIC (positive), 
OTHER_DISEASE (positive), OBESITY (positive), TOBACCO(negative), INTUBED (positive),
ICU(negative)

<br>  

**Run knn model**
```{r, include=FALSE}
#set.seed(2)
#knn_model <- train(data = Covid_train, DEATH~., method = "knn", metric = "Accuracy", 
      #tuneGrid = expand.grid(k=c(1, 5, 7)), preProcess = c("center", "scale"))
#knn_model
```
**Comment**<br>  
KNN model failed since there are too many ties and KNN can't deal with ties. This 
means that there are many similar data points which have the same distance.
<br>  

**Run NB model**
```{r, include = FALSE}
library(e1071)
set.seed(3)
nb_model <- naiveBayes(DEATH~., data = Covid_train)
nb_model
```
<br>  

**Run Decision Tree**
```{r}
set.seed(4)
#agnes or hclust object does not work with later prediction
dt_model <- rpart(DEATH~., data = Covid_train, method = "class") #class for binary
rpart.plot(dt_model, extra = 110, main = "Dendrogram of rpart") 
```

**Model Testing**
```{r}
#Test the logistic regression model and return in probability
log_test_prob <- predict(log_model, Covid_test, type = "response")

#Test the knn model
#knn_test_prob <- predict(knn_model, Covid_test, type = "prob")

#Test the nb model
nb_test_prob <- predict(nb_model, Covid_test, type = "raw")

#Test the dt model- (predict does not apply to "hclust" or "agnes" object)
dt_test_prob <- predict(dt_model, Covid_test, type = "prob")
```
<br>  
<br>  

**Model Comparison: Thresholding, best cutoff point, confusion table and ROC**
```{r}
#logistic regression 
pred_log_test <- prediction(log_test_prob, Covid_test$DEATH)#create prediction obj

#TPR FPR plot
roc_perf_log_test <- performance(pred_log_test, measure = "tpr", x.measure = "fpr")
plot(roc_perf_log_test,colorize=TRUE,print.cutoffs.at=seq(0.1,by=0.1)) 
#TPR/FPR cutoff graph<br>  
```

```{r}
#Logistic regression AUC value
auc.perf = performance(pred_log_test, measure = "auc")
auc.perf@y.values
```

```{r}
#Confusion table
confusionMatrix(as.factor(ifelse(log_test_prob>0.1, "1", "2")), Covid_test$DEATH, positive = "1")
```
<br>  

**Logistic Regression Metric**<br>  
True Positive (TP) = 13442 <br>  
True Negative (TN) = 264 <br>  
False Positive (FP) = 101037 <br>  
False Negative (FN) = 2812 <br>  
Miscalculations = 103849 <br>  
Accuracy = 11.66% <br>  
Sensitivity = 82.70% <br>  
Specificity = 0.26% <br>  
<br>  
<br>  

```{r}
#Naive Bayes 
pred_nb_test <- prediction(nb_test_prob[,1], Covid_test$DEATH)
roc_perf_nb_test <- performance(pred_nb_test, measure = "tpr", x.measure = "fpr")
plot(roc_perf_nb_test,colorize=TRUE,print.cutoffs.at=seq(0.1,by=0.1))
```
<br>  
```{r}
#Calculate ROC value for binary classifier
roc.curve(Covid_test$DEATH, nb_test_prob[,1], plotit= F)
```

```{r}
confusionMatrix(as.factor(ifelse(nb_test_prob[,1]>0.01, "1", "2")), Covid_test$DEATH, positive = "1")
```
<br>  
<br>  
**Naive Bayes Metric**<br>  
True Positive (TP) = 13857 <br>  
True Negative (TN) = 85337 <br>  
False Positive (FP) = 15964 <br>  
False Negative (FN) = 2397 <br>  
Miscalculations = 18361 <br>  
Accuracy = 84.38% <br>  
Sensitivity = 85.25% <br>  
Specificity = 84.24% <br>  
<br>  


```{r}
#decision tree (dt): create prediction object for ROCR evaluation 
pred_dt_test <- prediction(dt_test_prob[,1], Covid_test$DEATH)
roc_perf_dt_test <- performance(pred_dt_test, measure = "tpr", x.measure = "fpr")
plot(roc_perf_nb_test,colorize=TRUE,print.cutoffs.at=seq(0.1,by=0.1))
```

```{r}
#Calculate ROC value for binary classifier
roc.curve(Covid_test$DEATH, dt_test_prob[,1], plotit= F)
```

```{r}
confusionMatrix(as.factor(ifelse(dt_test_prob[,1]>0.2, "1", "2")), Covid_test$DEATH, positive = "1")
```
<br>  
<br>  

**Decision Tree Metric**<br>  
True Positive (TP) = 14752 <br>  
True Negative (TN) = 82584 <br>  
False Positive (FP) = 18717 <br>  
False Negative (FN) = 1502 <br>  
Miscalculations = 20219 <br>  
Accuracy = 82.8% <br>  
Sensitivity = 90.76% <br> 
Specificity = 81.52% <br>  
<br>  
<br>  

**Conclusion**<br>  
The model aims to reduce false negatives and tolerates more on false positives. 
It will cost more to miss a covid positive patient than to mis-classify a negative
one. <br>  
The next step is to discover the cost for false positive and false negative patient
to adjust the model to save the total cost. 
